---
title: "Gov 50 Final Project"
author: "Ebony McDonald"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to... voter turnout?

```{r}
library(haven)
data <- read_sav("Data/W110_Jun22/ATP W110.sav")
```

```{r}
library(tidyverse)
library(dplyr)
library(infer)
```




##
1 White
2 Black or African-American
3 Asian or Asian-American
4 Mixed Race
5 Or some other race
99 Don’t know/Refused

TRAITREPa Compared to other Americans, would you say Republicans are…
1 A lot more moral
2 Somewhat more moral
3 About the same
4 Somewhat more immoral
5 A lot more immoral

TRAITREPb Compared to other Americans, would you say Republicans are…
1 A lot more hard-working
2 Somewhat more hard-working
3 About the same
4 Somewhat more lazy
5 A lot more lazy


TRAITREPc Compared to other Americans, would you say Republicans are…
1 A lot more open-minded
2 Somewhat more open-minded
3 About the same
4 Somewhat more closed-minded
5 A lot more closed-minded

TRAITREPe Compared to other Americans, would you say Republicans are…
1 A lot more intelligent
2 Somewhat more intelligent
3 About the same
4 Somewhat more UNintelligent
5 A lot more UNintelligent

TRAITREPf Compared to other Americans, would you say Republicans are…
1 A lot more honest
2 Somewhat more honest
3 About the same
4 Somewhat more DIShonest
5 A lot more DIShonest




```




```{r}
polarity <- data |>
  filter(F_PARTY_FINAL == 1 | F_PARTY_FINAL == 2)

polarity <- polarity |>
  filter(TRAITREPa_W110 < 99)

polarity <- polarity |>
  filter(TRAITDEMa_W110 < 99)

polarity <- polarity |>
  mutate(
    Party = case_when(
      F_PARTY_FINAL == 1 ~ "Republican",
      F_PARTY_FINAL == 2 ~ "Democrat"))


character_chart <- polarity |>
  group_by(Party)|>
  summarize(moral_eval_rep = mean(TRAITREPa_W110),
            moral_eval_dem = mean(TRAITDEMa_W110),
            hardwork_eval_rep = mean(TRAITREPb_W110),
            hardwork_eval_dem = mean(TRAITDEMb_W110),
            openmind_eval_rep = mean(TRAITREPc_W110),
            openmind_eval_dem = mean(TRAITDEMc_W110),
            smart_eval_rep = mean(TRAITREPe_W110),
            smart_eval_dem = mean(TRAITDEMe_W110),
            honest_eval_rep = mean(TRAITREPf_W110),
            honest_eval_dem = mean(TRAITDEMf_W110))




  

```





